---
title: "2024-11-22_analysis"
author: "01schau"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Importing data saved as rds (r file format)

```{r}
library(tidyverse)
```

```{r}
here::here("data/human_data_selection_dedup.rds")
human_selection <- readRDS(here::here("data/human_data_selection_dedup.rds"))
```

connection to the database in whonet (I made a connection to an old sqlite datafile - should make a conection to the original file that will be used), for that we need to install a new package:

Need to install RSQLite first

```{r}
#install.packages("RSQLite")
library("RSQLite")
```

```{r}
sqlite_path <- "C:/WHONET/Data/TZA-INIKA_TZ-2024.sqlite"
sqlite_path
dbconn <- DBI::dbConnect(RSQLite::SQLite(), 
                    here::here("C:/WHONET/Data/", "TZA-INIKA_TZ-2024.sqlite"))

str(dbconn)
print(dbconn)
```

or try this that gives the same output

```{r}
sqlite_path <- "C:/WHONET/Data/TZA-INIKA_TZ-2024.sqlite"
sqlite_path
dbconn <- DBI::dbConnect(
  RSQLite::SQLite(), 
  sqlite_path)

str(dbconn)
print(dbconn)
```

dbconn builds the connection

```{r}
dbconn
```

# list the tables in whonet

```{r}
dbListTables(dbconn)
```

read the data from the table (mydata is a recipe)

```{r}
mydata <- tbl(dbconn, "Isolates")
```

```{r}
mydata_recipe <- tbl(dbconn, "Isolates")
str(mydata_recipe)
head(mydata_recipe)
```

str look at the structure of the table? / head ask for output
we want to get all the data

```{r}
show_query(mydata_recipe)
```

output from the command:
show_query(mydata_recipe)
<SQL>
SELECT * the asterisk means all after
FROM `Isolates`

```{r}
mydata <- mydata_recipe %>% 
  collect()
```

We now have to datasets/-tables in R memory and now we want to combine them
There are different ways to join tables; inner-join, full-join, left-join, etc. If you do not know what you want, use the help function. Often we use the left-join.

```{r}
#my_innerjoin <-
  human_selection %>%
  # selecting few columns for testing
  # this can be used to do a short selection of the columns if not all are required
  select(1:3) %>% 
  dplyr::inner_join(mydata, by = c("INIKA_OH_TZ_ID" = "PATIENT_ID")) %>% 
  head() %>% 
  view()

my_innerjoin %>%
  str()
```

Would this work? Something with the order, no this will not work in the same way - :

```{r}
mydata %>% 
  dplyr::inner_join(
    human_selection %>% 
      select(1:3), 
    by = c("PATIENT_ID" = "INIKA_OH_TZ_ID")) %>% 
      head() %>% 
      view()
```

Something is not working when I do this, I get an error and can not figure out why - yet

