---
title: "2024-11-18_course-part1"
author: "01schau"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Course part 1 - Laboratory

-   We selected and sub-cultured 20 isolates from the -80 freezer
    -   8 presumptive *Salmonella* spp. isolates
    -   6 presumptive *E. coli* isolates
    -   6 presumptive *Klebsiella* spp. isolates

# Installation of packages

```{r}
install.packages("here")
install.packages("DiagrammeR")
```

# ctrl + alt + I 
Gives the start and stop of a code/ a cell (cheat command, short code)

```{r}
angry_or_not <- "not"
```

To give the content to an object
```{r}
angry_or_not
```

Madelaine, are you angry or not? `r print(angry_or_not)`

```{r}
"Madelaine, are you angry or not?" 
`r print(angry_or_not)`
```

# Working with your data

```{r}
here::here()
```
```{r}
path_data <- here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")

path_data
```

```{r}
human_data <-
  readr::read_csv2(path_data)
```

# Wednesday - day 3

## Packages into memory

```{r}
library(here)
library(tidyverse)
```

## Getting help with packages and what does a function do

F1

```{r}
?library()
help("library")
```

## we get our data

```{r}
path_data <- here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")

path_data
```

```{r}
human_data <-
  readr::read_csv2(path_data)
```

## Exploration of a data fram

- Names of the columns
```{r}
colnames(human_data)
```

- getting the column names into a vector for pasting and manipulations
- i can use a function within a function
```{r}
dput(
  colnames(human_data)
  )
```

- explanation: c() is to make vectors
```{r}
test <- c("one", "2", 3)
test
typeof(test)
class(test)
```
- figure out if the character is numeric or text or something else
- what type of objects do we have?

# sign for an assignment: <-

```{r}
test <- c("one", "2", 3)
test
test[3]
typeof(test[3])

class(test)
```
All the elements of a vector are of the same `type`(text, double = real, integer, date ....)

```{r}
test2 <- c(1,2,3)
typeof(test2)
class(test2)
test2*test2
```

```{r}
test2 * c(1,2)

```



